National Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Beliefs

### 2020-11-21 Analysis.Rmd
small changes

```{r global_options, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(fig.path="Figs/", 
                      message=FALSE, 
                      warning=FALSE,
                      collapse =TRUE,
                      echo=TRUE, 
                      #results="hide", 
                      fig.width= 10,
                      fig.height=8,
                      tidy = "styler")
```
```{r load_libraries, echo=FALSE, include=FALSE}
library("styler")
library("tidyverse")
library("lme4")
library("ggplot2")
library("tidyr")
library("ggeffects")
library("Amelia")
library("table1")
library("patchwork")
library("easystats")
library("parameters")
library("prettycode")
library("table1")
library("gghighlight")
library("sjPlot")
library("texreg")
library("see")
#library("viridis")
#library("ggpubr")
#library("wesanderson")
library("ggsci")
library("papaja")
#library("here")
library("citr")

```

@@

Processing of data.
```{r data processing, cache = TRUE, dependson = "ldf.5"}
# filter only those who are measured 
dfben  <-ldf.5 %>%
  dplyr:: filter(Wave !=2009)%>% # exclude wave 1 (no spirit measures that wave)
  dplyr::filter(YearMeasured==1)%>%
  dplyr::group_by(Id) %>% filter(n() > 2)%>% # select those who have responded to at least 3 waves # Also works for 5 and 7
  dplyr::filter(n() !=0)%>%
  dplyr::ungroup(Id)%>%
  dplyr:: mutate(Years = as.numeric(years))%>%
  dplyr:: mutate(yearsC = scale(years, center=TRUE,scale=FALSE))%>%
  dplyr::mutate(Age.10yrs = (Age/10))%>%
  dplyr::mutate(Religion.Church.Log = log(Religion.Church+ 1))%>%
  dplyr::mutate(Pol.Orient.S = scale(Pol.Orient, center = TRUE, scale = TRUE),
         Age.10yrs.C = scale(Age.10yrs, center = TRUE, scale = FALSE),
         Male = factor(Gender),
         Employed = factor(Employed),
         Partner = factor(Partner),
         Edu.S = scale(as.numeric(Edu) ,scale=TRUE,center=TRUE),
         Education = as.numeric(Edu),
         Urban = factor(Urban),
         Believe.Spirit = as.factor(Believe.Spirit),
         Believe.God = as.factor(Believe.God),
         EthnicCategories = factor(EthnicCats),
         LIFESAT.S = scale(LIFESAT, center = TRUE, scale = TRUE),
         PWI.S = scale(PWI, center = TRUE, scale = TRUE),
         Deprivation.S = scale(NZdep, scale=TRUE, center=TRUE),
         Relid.C = scale(Relid, scale = TRUE,center=TRUE),
         Religious = as.factor(Religious),
         Religion.Church.Log.C = scale(Religion.Church.Log, center = TRUE, scale = FALSE))
```
```{r}
# nice labels
table1::label(dfben$Age.10yrs.C) <- "Age in Decades (C)"
table1::label(dfben$PWI) <- "Personal Wellbeing"
table1::label(dfben$LIFESAT) <- "Life Satisfaction"

@@

Table of belief types
```{r}
table(dfben$Beliefs)
range(dfben$years)
range(dfben$Years)
```

@@

```{r}
table1::table1(~ Beliefs|Wave, data = dfben, overall = T)
```

```{r, include =FALSE}
#furniture::table1(dfben, Age, NZdep,Education,Employed,EthnicCategories,Male,Partner, Pol.Orient,Urban, Beliefs, LIFESAT,PWI,
                  #  splitby = ~Wave, overall=F, output = "latex2", booktabs=TRUE)
```

<!-- ```{r, include=FALSE} -->
<!-- r<-easystats::report_participants( -->
<!--   dfben, -->
<!--   age = "Age", -->
<!--   sex = "Male", -->
<!--   education = "Education", -->
<!--   participants = "Id", -->
<!--   group = "Wave", -->
<!--   spell_n = FALSE, -->
<!-- ) -->
<!-- ??report_participants -->
<!-- ``` -->

@@

```{r check collinearity}
# Note High VIFS from 
performance::check_collinearity(mod.1)
performance::check_collinearity(mod.2)
```

Inference holds but VIFS years x beliefs when years are centered.

```{r model centering years}
library(lme4)
mod.1a <-lmer (PWI ~  1 +  yearsC * Beliefs +  Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben, na.action = na.omit)
mod.2a <-lmer (LIFESAT ~ 1 + yearsC * Beliefs+ Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben, na.action = na.omit)
summary(mod.1a)
summary(mod.2a)
# identical result
performance::check_collinearity(mod.1a)
performance::check_collinearity(mod.2a)

# address reviewer worry about deprivation and partner being correlated
m<-lmer (NZdep ~ Partner + (1|Id), data= dfben, na.action = na.omit)
tab_model(m)
model_parameters(m)
```

Model is robust without covariates

```{r robust to years and beliefs only}
library(lme4)
mod.1aa <-lmer (PWI ~  1 +  years * Beliefs + (1|Id), data= dfben, na.action = na.omit)
mod.2aa <-lmer (LIFESAT ~ 1 + years * Beliefs + (1|Id), data= dfben, na.action = na.omit)
)
# identical result
sjPlot::tab_model(mod.1aa,mod.2aa)
```

@@

```{r USE_2_coefficientplots}
g1/g2
range(dfben$yearsC)
```

```{r cache = TRUE, dependson = "mod.1"}
#### Predicted Effects ####
pr.PWI <- ggpredict(model = mod.1, terms = c("Years [0:9]","Beliefs"), 
                    ci.lvl = 0.95,
                    type = "fe",
                    typical = "mean",
                   # condition = c(Euro =1),
                    back.transform = TRUE,
                    ppd = FALSE,
                    interval = "confidence")
p1<-plot(pr.PWI, facets = T) + 
 # scale_y_continuous(limits=c(4.75,5.25) )  + 
  gghighlight()  +  theme_blank()+ 
  ggtitle("Predicted Values of Personal Wellbeing") 
p1
pr.PWI
```
```{r pwi_expected}
```

@@

<!-- ```{r eq1} -->
<!-- library(equatiomatic) -->
<!-- print(extract_eq(mod.1, wrap = TRUE,terms_per_line =3)) -->
<!-- ``` -->

<!-- PWI Model 2 equation -->
<!-- ```{r eq2} -->
<!-- library(equatiomatic) -->
<!-- print(extract_eq(mod.2, wrap = TRUE, use_coefs=TRUE,terms_per_line = 3)) -->
<!-- ``` -->

<!-- ## report -->

<!-- ### Model 1 -->
<!-- model 1 report: for publication change all "significants" to "reliable" -->

<!-- short report: -->
<!-- ```{r, report1, cache =TRUE, dependson=mod.1,include=FALSE} -->
<!-- easystats::model_text(mod.1) -->
<!-- ``` -->

<!-- short report -->
<!-- ```{r, cache=TRUE, include=FALSE } -->
<!-- library(report) -->
<!-- r1<- easystats::report(mod.1) -->
<!-- easystats::table_short(r1) -->
<!-- easystats::text_short(r1) -->
<!-- ``` -->

<!-- ```{r, cache =TRUE, dependson=mod.1, include=FALSE} -->
<!-- r1<- easystats::report(mod.1) -->
<!-- easystats::text_long(r1) -->
<!-- ``` -->
<!-- short report -->
<!-- ```{r, cache =TRUE, dependson=mod.2,include=FALSE} -->
<!-- # Model 2 -->
<!-- # For publication change all "significants" to "reliable" -->

<!-- easystats::model_text(mod.2) -->
<!-- ``` -->

<!-- short report -->
<!-- ```{r,include=FALSE} -->
<!-- r2 <- easystats::report(mod.2) -->
<!-- easystats::text_short(r2) -->
<!-- easystats::table_short(r2) -->
<!-- ``` -->

<!-- long report -->
<!-- ```{r,include=FALSE} -->
<!-- r2<- easystats::report(mod.2) -->
<!-- easystats::text_long(r2) -->
<!-- ``` -->

@@

```{r predicted_lifesat_imputed, cache=TRUE}
explectationGraphLS
```
```{r combograph_imputed}
# Combo_Graph
#explectationGraphPWI/explectationGraphLS + plot_annotation(title = "Imputed DataSet")
##
```

>>>>

```{r global_options, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(fig.path="Figs/", 
                      message=FALSE, 
                      warning=FALSE,
                      collapse =TRUE,
                      echo=TRUE, 
                      #results="hide", 
                      fig.width= 10,
                      fig.height=8,
                      tidy = "styler")
```
```{r load_libraries, echo=FALSE, include=FALSE}
library("styler")
library("tidyverse")
library("lme4")
library("ggplot2")
library("tidyr")
library("ggeffects")
library("Amelia")
library("table1")
library("patchwork")
library("parameters")
library("prettycode")
library("table1")
library("gghighlight")
library("sjPlot")
library("see")
#library("viridis")
#library("ggpubr")
#library("wesanderson")
library("ggsci")
library("papaja")
#library("here")
library("citr")
library("texreg")
```

@@

Processing of data.
```{r data processing, cache = TRUE, dependson = "ldf.5"}
# filter only those who are measured 
dfben  <-ldf.5 %>%
  dplyr:: filter(Wave !=2009)%>% # exclude wave 1 (no spirit measures that wave)
  dplyr::filter(YearMeasured==1)%>%
  dplyr::group_by(Id) %>% filter(n() > 2)%>% # select those who have responded to at least 3 waves # Also works for 5 and 7
  dplyr::filter(n() !=0)%>%
  dplyr::ungroup(Id)%>%
  dplyr::mutate(Years = as.numeric(years))%>%
  dplyr::mutate(Age.10yrs = (Age/10))%>%
  dplyr::mutate(Religion.Church.Log = log(Religion.Church+ 1))%>%
  dplyr::mutate(Pol.Orient.S = scale(Pol.Orient, center = TRUE, scale = TRUE),
         Age.10yrs.C = scale(Age.10yrs, center = TRUE, scale = FALSE),
         Male = factor(Gender),
         Employed = factor(Employed),
         Partner = factor(Partner),
         Edu.S = scale(as.numeric(Edu) ,scale=TRUE,center=TRUE),
         Education = as.numeric(Edu),
         Urban = factor(Urban),
         Believe.Spirit = as.factor(Believe.Spirit),
         Believe.God = as.factor(Believe.God),
         EthnicCategories = factor(EthnicCats),
         LIFESAT.S = scale(LIFESAT, center = TRUE, scale = TRUE),
         PWI.S = scale(PWI, center = TRUE, scale = TRUE),
         Deprivation.S = scale(NZdep, scale=TRUE, center=TRUE),
         Relid.C = scale(Relid, scale = TRUE,center=TRUE),
         Religious = as.factor(Religious),
         Religion.Church.Log.C = scale(Religion.Church.Log, center = TRUE, scale = FALSE))

# nice labels
table1::label(dfben$Age.10yrs.C) <- "Age in Decades (C)"
table1::label(dfben$PWI) <- "Personal Wellbeing"
table1::label(dfben$LIFESAT) <- "Life Satisfaction"

@@

Table of belief types
```{r}
table(dfben$Beliefs)
```

@@

```{r, include =FALSE}
furniture::table1(dfben, Age, NZdep,Education,Employed,EthnicCategories,Male,Partner, Pol.Orient,Urban, Beliefs, LIFESAT,PWI,
                    splitby = ~Wave, overall=F, output = "latex2", booktabs=TRUE)
```

```{r, include=FALSE}
r<-report::report_participants(
  dfben,
  age = "Age",
  sex = "Male",
  education = "Education",
  participants = "Id",
  group = "Wave",
  spell_n = FALSE,
)
```

@@

@@

```{r USE_2_coefficientplots}
g1/g2
```

```{r cache = TRUE, dependson = "mod.1"}
#### Predicted Effects ####
pr.PWI <- ggpredict(model = mod.1, terms = c("Years [0:9]","Beliefs"), 
                    ci.lvl = 0.95,
                    type = "fe",
                    typical = "mean",
                   # condition = c(Euro =1),
                    back.transform = TRUE,
                    ppd = FALSE,
                    interval = "confidence")
p1<-plot(pr.PWI, facets = T) + 
 # scale_y_continuous(limits=c(4.75,5.25) )  + 
  gghighlight()  +  theme_blank()+ 
  ggtitle("Predicted Values of Personal Wellbeing")

pr.PWI
```
```{r pwi_expected}
```

@@

```{r eq1}
library(equatiomatic)
print(extract_eq(mod.1, wrap = TRUE,terms_per_line =3))
```

PWI Model 2 equation
```{r eq2}
library(equatiomatic)
print(extract_eq(mod.2, wrap = TRUE, use_coefs=TRUE,terms_per_line = 3))
```

## report

### Model 1
model 1 report: for publication change all "significants" to "reliable"

short report:
```{r, report1, cache =TRUE, dependson=mod.1,include=FALSE}
report::model_text(mod.1)
```

short report
```{r, cache=TRUE, include=FALSE }
library(report)
r1<- report::report(mod.1)
table_short(r1)
text_short(r1)
```

```{r, cache =TRUE, dependson=mod.1, include=FALSE}
r1<- report::report(mod.1)
report::text_long(r1)
```
short report
```{r, cache =TRUE, dependson=mod.2,include=FALSE}
# Model 2
# For publication change all "significants" to "reliable"

report::model_text(mod.2)
```

short report
```{r,include=FALSE}
r2 <- report::report(mod.2)
report::text_short(r2)
report::table_short(r2)
```

long report
```{r,include=FALSE}
r2<- report::report(mod.2)
report::text_long(r2)
```

@@

```{r predicted_lifesat_imputed, cache=TRUE}
explectationGraphLS
```
```{r combograph_imputed}
# Combo_Graph
#explectationGraphPWI/explectationGraphLS + plot_annotation(title = "Imputed DataSet")
##
sessionInfo()
```
